A conditional permutation-based approach to test confounder and center effects in machine learning models

### 2021-10-21 simulated/normality_and_linearity_violation.tex
example notebook

\begin{lstlisting}[language=Python]
import numpy as np
import seaborn as sns
import matplotlib.pyplot as plt
from sklearn.linear_model import Ridge
sns.set(rc={"figure.figsize":(6, 2)})
sns.set_style("white")
\end{lstlisting}

@@

\begin{lstlisting}[language=Python]
n=300
p=5
rng = np.random.default_rng(42)
y = np.arange(n)
y = (y - y.mean())/y.std()

X_true = np.repeat(y, p).reshape(n,p)
for i in range(1,4):
    X_true[:,i] = np.tanh(X_true[:,i]*2)

X=X_true + rng.normal(0,0.1, (n,p))

for i in range(4):
    sns.lineplot(x=y, y=X_true[:,i])
    sns.scatterplot(x=y, y=X[:,i])
    plt.show()
\end{lstlisting}

@@

\begin{lstlisting}[language=Python]
model = Ridge(alpha=0)
model.fit(y=y, X=X)
yhat = model.predict(X)
sns.scatterplot(x=y, y=yhat, alpha=0.3)
#sns.lineplot(x=y, y=y, linestyle="--")
model.coef_
\end{lstlisting}

\begin{lstlisting}
array([ 0.47902429,  0.00946499,  0.07378478, -0.02828672,  0.47070535])
\end{lstlisting}

@@

\begin{lstlisting}[language=Python]
model = Ridge(alpha=2000)
model.fit(y=y, X=X)
yhat = model.predict(X)
sns.scatterplot(x=y, y=yhat, alpha=0.3)
#sns.lineplot(x=y, y=y, linestyle="--")
model.coef_
\end{lstlisting}

\begin{lstlisting}
array([0.09472368, 0.07472972, 0.07338482, 0.07440682, 0.09463765])
\end{lstlisting}

@@

\begin{lstlisting}[language=Python]
\end{lstlisting}

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\ImportTok{import}\NormalTok{ numpy }\ImportTok{as}\NormalTok{ np}
\ImportTok{import}\NormalTok{ seaborn }\ImportTok{as}\NormalTok{ sns}
\ImportTok{import}\NormalTok{ matplotlib.pyplot }\ImportTok{as}\NormalTok{ plt}
\ImportTok{from}\NormalTok{ sklearn.linear_model }\ImportTok{import}\NormalTok{ Ridge}
\NormalTok{sns.}\BuiltInTok{set}\NormalTok{(rc}\OperatorTok{=}\NormalTok{\{}\StringTok{"figure.figsize"}\NormalTok{:(}\DecValTok{6}\NormalTok{, }\DecValTok{2}\NormalTok{)\})}
\NormalTok{sns.set_style(}\StringTok{"white"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n}\OperatorTok{=}\DecValTok{300}
\NormalTok{p}\OperatorTok{=}\DecValTok{5}
\NormalTok{rng }\OperatorTok{=}\NormalTok{ np.random.default_rng(}\DecValTok{42}\NormalTok{)}
\NormalTok{y }\OperatorTok{=}\NormalTok{ np.arange(n)}
\NormalTok{y }\OperatorTok{=}\NormalTok{ (y }\OperatorTok{-}\NormalTok{ y.mean())}\OperatorTok{/}\NormalTok{y.std()}

\NormalTok{X_true }\OperatorTok{=}\NormalTok{ np.repeat(y, p).reshape(n,p)}
\ControlFlowTok{for}\NormalTok{ i }\KeywordTok{in} \BuiltInTok{range}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{):}
\NormalTok{    X_true[:,i] }\OperatorTok{=}\NormalTok{ np.tanh(X_true[:,i]}\OperatorTok{*}\DecValTok{2}\NormalTok{)}

\NormalTok{X}\OperatorTok{=}\NormalTok{X_true }\OperatorTok{+}\NormalTok{ rng.normal(}\DecValTok{0}\NormalTok{,}\FloatTok{0.1}\NormalTok{, (n,p))}

\ControlFlowTok{for}\NormalTok{ i }\KeywordTok{in} \BuiltInTok{range}\NormalTok{(}\DecValTok{4}\NormalTok{):}
\NormalTok{    sns.lineplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{X_true[:,i])}
\NormalTok{    sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{X[:,i])}
\NormalTok{    plt.show()}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model }\OperatorTok{=}\NormalTok{ Ridge(alpha}\OperatorTok{=}\DecValTok{0}\NormalTok{)}
\NormalTok{model.fit(y}\OperatorTok{=}\NormalTok{y, X}\OperatorTok{=}\NormalTok{X)}
\NormalTok{yhat }\OperatorTok{=}\NormalTok{ model.predict(X)}
\NormalTok{sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{yhat, alpha}\OperatorTok{=}\FloatTok{0.3}\NormalTok{)}
\CommentTok{#sns.lineplot(x=y, y=y, linestyle="--")}
\NormalTok{model.coef_}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
array([ 0.47902429,  0.00946499,  0.07378478, -0.02828672,  0.47070535])
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{model }\OperatorTok{=}\NormalTok{ Ridge(alpha}\OperatorTok{=}\DecValTok{2000}\NormalTok{)}
\NormalTok{model.fit(y}\OperatorTok{=}\NormalTok{y, X}\OperatorTok{=}\NormalTok{X)}
\NormalTok{yhat }\OperatorTok{=}\NormalTok{ model.predict(X)}
\NormalTok{sns.scatterplot(x}\OperatorTok{=}\NormalTok{y, y}\OperatorTok{=}\NormalTok{yhat, alpha}\OperatorTok{=}\FloatTok{0.3}\NormalTok{)}
\CommentTok{#sns.lineplot(x=y, y=y, linestyle="--")}
\NormalTok{model.coef_}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
array([0.09472368, 0.07472972, 0.07338482, 0.07440682, 0.09463765])
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]

\end{Highlighting}
\end{Shaded}
